---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Olivia Zhang"
date: "02/23/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r filter-dn-ak}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are `r nrow(dn_ak)` Denny’s locations in Alaska.

```{r filter-lq-ak}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are `r nrow(lq_ak)` La Quinta locations in Alaska.

### Exercise 2

```{r join}
dn_ak$establishment <- "dn"
lq_ak$establishment <- "lq"

dn_lq_ak1 <- 
  full_join(dn_ak, lq_ak, by = join_by(address, city, state, zip, longitude, latitude, establishment))
#I did this before I see your notes. My way of joining the two data frames allows the same variables for two locations, while your way of joining datasets allows for pairing and later including distance information into the joined dataset.
```

```{r dn1-distance}
#dn1
dn1_lq1 <- sqrt((dn_lq_ak1$longitude[1] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[1] - dn_lq_ak1$latitude[4])^2)
dn1_lq2 <- sqrt((dn_lq_ak1$longitude[1] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[1] - dn_lq_ak1$latitude[5])^2)

#dn2
dn2_lq1 <- sqrt((dn_lq_ak1$longitude[2] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[2] - dn_lq_ak1$latitude[4])^2)
dn2_lq2 <- sqrt((dn_lq_ak1$longitude[2] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[2] - dn_lq_ak1$latitude[5])^2)

#dn3
dn3_lq1 <- sqrt((dn_lq_ak1$longitude[3] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[3] - dn_lq_ak1$latitude[4])^2)
dn3_lq2 <- sqrt((dn_lq_ak1$longitude[3] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[3] - dn_lq_ak1$latitude[5])^2)
#I only noticed your notes after I typed in all these. I didn't want to delete it so here they are. They are not used for later calculations or visualizations.
```
There are six pairings. 

### Exercise 3

Well, I guess I did in my own way above. I'll follow the steps now. 
```{r join-mason}
dn_lq_ak <- full_join(dn_ak, lq_ak, 
                      by = "state")
dn_lq_ak
```


### Exercise 4
There are `r nrow(dn_lq_ak)` observations in the joined dn_lq_ak data frame. The names are `r names(dn_lq_ak)`.

### Exercise 5

```{r haversine, echo=FALSE}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```


### Exercise 6

```{r distance}
dn_lq_ak$distance <- haversine(dn_lq_ak$longitude.x, dn_lq_ak$latitude.x, dn_lq_ak$longitude.y, dn_lq_ak$latitude.y, round = 3)
```

### Exercise 7

```{r min-distance}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8

```{r ak-vis}
dn_lq_ak %>%
  filter(dn_lq_ak$distance %in% dn_lq_ak_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in Alaska",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```

As shown in the graph, in Alaska, La Quinta locations are all near Denny’s. The distance between the nearest Denny's and La Quinta are `r dn_lq_ak_mindist$closest`.

### Exercise 9

```{r nc}
#filter out NC
dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

#join the data frames
dn_nc$establishment <- "dn"
lq_nc$establishment <- "lq"

dn_lq_nc <- full_join(dn_nc, lq_nc, 
                      by = "state")
dn_lq_nc

#calculate distances for all pairings 
dn_lq_nc$distance <- haversine(dn_lq_nc$longitude.x, dn_lq_nc$latitude.x, dn_lq_nc$longitude.y, dn_lq_nc$latitude.y, round = 3)

#find the minimum distance
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#visualize the shortest distances
dn_lq_nc %>%
  filter(dn_lq_nc$distance %in% dn_lq_nc_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in North Carolina",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```

### Exercise 10

```{r tx}
#filter out TX
dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

#join the data frames
dn_tx$establishment <- "dn"
lq_tx$establishment <- "lq"

dn_lq_tx <- full_join(dn_tx, lq_tx, 
                      by = "state")
dn_lq_tx

#calculate distances for all pairings 
dn_lq_tx$distance <- haversine(dn_lq_tx$longitude.x, dn_lq_tx$latitude.x, dn_lq_tx$longitude.y, dn_lq_tx$latitude.y, round = 3)

#find the minimum distance
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#visualize the shortest distances
dn_lq_tx %>%
  filter(dn_lq_tx$distance %in% dn_lq_tx_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x, 
    alpha = .5
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y, 
    alpha = .5
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in Texas",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     ) +
  guides(alpha = FALSE)
```

### Exercise 11

```{r ca}
#filter out CA
dn_ca <- dennys %>%
  filter(state == "CA")

lq_ca <- laquinta %>%
  filter(state == "CA")

#join the data frames
dn_ca$establishment <- "dn"
lq_ca$establishment <- "lq"

dn_lq_ca <- full_join(dn_ca, lq_ca, 
                      by = "state")
dn_lq_ca

#calculate distances for all pairings 
dn_lq_ca$distance <- haversine(dn_lq_ca$longitude.x, dn_lq_ca$latitude.x, dn_lq_ca$longitude.y, dn_lq_ca$latitude.y, round = 3)

#find the minimum distance
dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#visualize the shortest distances
dn_lq_ca %>%
  filter(dn_lq_ca$distance %in% dn_lq_ca_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x,
    alpha = .5
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y,
    alpha = .5
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in California",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     ) +
  guides(alpha = FALSE)
```


### Exercise 12

Mitch Hedberg's joke about La Quinta is the Spanish for next for Denny's is true for most state except for North Carolina, where there are two La Quinta's that is still pretty far from the nearest Denny's. It is worth noting that it is generally true that if you find a La Quinta location, there is a Denny's near it, but this joke doesn't garantee that if you find a Denny's location, a La Quinta's location is necessarily nearby.



