---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Olivia Zhang"
date: "02/23/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r filter-dn-ak}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are `r nrow(dn_ak)` Denny’s locations in Alaska.

```{r filter-lq-ak}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are `r nrow(lq_ak)` La Quinta locations in Alaska.

### Exercise 2

```{r join}
dn_ak$establishment <- "dn"
lq_ak$establishment <- "lq"

dn_lq_ak1 <- 
  full_join(dn_ak, lq_ak, by = join_by(address, city, state, zip, longitude, latitude, establishment))
```

```{r dn1-distance}
#dn1
dn1_lq1 <- sqrt((dn_lq_ak1$longitude[1] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[1] - dn_lq_ak1$latitude[4])^2)
dn1_lq2 <- sqrt((dn_lq_ak1$longitude[1] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[1] - dn_lq_ak1$latitude[5])^2)

#dn2
dn2_lq1 <- sqrt((dn_lq_ak1$longitude[2] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[2] - dn_lq_ak1$latitude[4])^2)
dn2_lq2 <- sqrt((dn_lq_ak1$longitude[2] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[2] - dn_lq_ak1$latitude[5])^2)

#dn3
dn3_lq1 <- sqrt((dn_lq_ak1$longitude[3] - dn_lq_ak1$longitude[4])^2 + (dn_lq_ak1$latitude[3] - dn_lq_ak1$latitude[4])^2)
dn3_lq2 <- sqrt((dn_lq_ak1$longitude[3] - dn_lq_ak1$longitude[5])^2 + (dn_lq_ak1$latitude[3] - dn_lq_ak1$latitude[5])^2)
```
There are six pairings. 

### Exercise 3

Well, I guess I did in my own way above. I'll follow the steps now. 
```{r join-mason}
dn_lq_ak <- full_join(dn_ak, lq_ak, 
                      by = "state")
dn_lq_ak
```


### Exercise 4
There are `r nrow(dn_lq_ak)` observations in the joined dn_lq_ak data frame. The names are `r names(dn_lq_ak)`.

### Exercise 5

```{r haversine, echo=FALSE}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```


### Exercise 6

```{r distance}
dn_lq_ak$distance <- haversine(dn_lq_ak$longitude.x, dn_lq_ak$latitude.x, dn_lq_ak$longitude.y, dn_lq_ak$latitude.y, round = 3)
```

### Exercise 7

```{r min-distance}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8

```{r ak-vis}
dn_lq_ak %>%
  filter(dn_lq_ak$distance %in% dn_lq_ak_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in Alaska",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```

As shown in the graph, in Alaska, La Quinta locations are all near Denny’s. The distance between the nearest Denny's and La Quinta are `r dn_lq_ak_mindist$closest`.

### Exercise 9

```{r nc}
#filter out NC
dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

#join the data frames
dn_nc$establishment <- "dn"
lq_nc$establishment <- "lq"

dn_lq_nc <- full_join(dn_nc, lq_nc, 
                      by = "state")
dn_lq_nc

#calculate distances for all pairings 
dn_lq_nc$distance <- haversine(dn_lq_nc$longitude.x, dn_lq_nc$latitude.x, dn_lq_nc$longitude.y, dn_lq_nc$latitude.y, round = 3)

#find the minimum distance
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#visualize the shortest distances
dn_lq_nc %>%
  filter(dn_lq_nc$distance %in% dn_lq_nc_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in North Carolina",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```

### Exercise 10

```{r tx}
#filter out TX
dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

#join the data frames
dn_tx$establishment <- "dn"
lq_tx$establishment <- "lq"

dn_lq_tx <- full_join(dn_tx, lq_tx, 
                      by = "state")
dn_lq_tx

#calculate distances for all pairings 
dn_lq_tx$distance <- haversine(dn_lq_tx$longitude.x, dn_lq_tx$latitude.x, dn_lq_tx$longitude.y, dn_lq_tx$latitude.y, round = 3)

#find the minimum distance
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#visualize the shortest distances
dn_lq_tx %>%
  filter(dn_lq_tx$distance %in% dn_lq_tx_mindist$closest) %>%
  ggplot() +
  geom_point(mapping = aes(
    x = longitude.x,
    y = latitude.x,
    color = establishment.x
    )) +
  geom_point(mapping = aes(
    x = longitude.y,
    y = latitude.y,
    color = establishment.y
    )) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in Texas",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```


### Exercise 11

### Exercise 12





